---
title: "Assignment 2"
author: "Xinyi Lin"
date: "10/14/2019"
output:
  word_document: default
  html_document: default
---

```{r,message=FALSE}
library(tidyverse)
```

## Question 1

Let trueDLT = true DLT probability, sp = selection probability, anTreat = average number of patients treated, anDLT = average number of patients with DLT. The table of the 3+3 design is shown below

```{r}
# Question 1
get_res = function(i,p){
  res = rbinom(3,1,p)
  res = rbind(rep(i,3),res)
  return(res)
}
dose1 = rbind(c(1:5), c(0.017, 0.043, 0.10, 0.22, 0.41))
```

```{r, eval=FALSE, include=FALSE}
dose = dose1
# one simulation
set.seed(123)
j = 1
res = get_res(dose[1,j], dose[2,j])
trial = res
t_v = rep(NA,30)
i = 0
while (sum(res[2,])<=1) {
    i = i+1
    t_v[i] = j
  if (sum(res[2,])==0){
    j = j+1
    res = get_res(dose[1,j], dose[2,j])
    trial = cbind(trial, res)
  } else if (sum(res[2,])==1 & t_v[i-1]!=j){
    j = j
    res = cbind(res, res = get_res(dose[1,j], dose[2,j]))
  } else if (sum(res[2,])==1 & t_v[i-1]==j){
    j = j+1
    res = get_res(dose[1,j], dose[2,j])
    trial = cbind(trial, res)
  }
}
trial = cbind(trial,res)
trial
```

```{r, eval=FALSE, include=FALSE}
dose = dose1
trail = c(0,0)
# one simulation
for (i in 1:5) {
  res = get_res(dose[1,i], dose[2,i])
  trail = cbind(trail, res)
  if (sum(res[2,])>=2) {
    mtd = i-1
    break
    } else if(sum(res[2,]==1)){
      ext_res = get_res(dose[1,i], dose[2,i])
      trail= cbind(trail,ext_res)
      res = cbind(res, ext_res)
    if(sum(res[2,])>=2) {
      mtd = i-1
      break
    }
    }
  mtd = i
}
trail[,-1]
mtd
```

```{r}
sim3p3 = function(dose, N = 1000){
  mtd_v = rep(NA,1000)
  com_trail = c(0,0)
  for (j in 1:N) {
    mtd = 0
    trail = c(0,0)
    for (i in 1:5) {
      res = get_res(dose[1,i], dose[2,i])
      trail = cbind(trail, res)
      if (sum(res[2,])>=2) {
        mtd = i-1
        break
        } else if(sum(res[2,]==1)){
          ext_res = get_res(dose[1,i], dose[2,i])
          trail= cbind(trail,ext_res)
          res = cbind(res, ext_res)
          if(sum(res[2,])>=2) {
            mtd = i-1
            break
          }
        }
      mtd = i
    }
    mtd_v[j] = mtd
    com_trail = cbind(com_trail, trail[,-1])
  }
  trail_res = t(com_trail[,-1])
  colnames(trail_res) = c("dose", "res")
  return(list(mtd = mtd_v, trail_res = trail_res))
}
```

```{r}
show_table = function(sim3p3, dose){
  sp = table(sim3p3$mtd)
  per_m = sim3p3$trail_res %>% 
    as.data.frame() %>% 
    group_by(dose) %>% 
    summarize(n = n(),
              DLT = sum(res)) %>% 
    mutate(anTreat = n/1000,
           anDLT = DLT/1000,
           s = as.numeric(sp[c(2:6)]),
           sp = s/1000) 
  per_m$trueDLT = as.numeric(dose[2,])
  per_m %>% 
    select(dose, trueDLT, sp, anTreat, anDLT) %>% 
    knitr::kable()
}
```

```{r, warning=FALSE}
set.seed(123)
sim3p31 = sim3p3(dose1)
show_table(sim3p31,dose1)
```

## Question 2

```{r}
set.seed(123)
# a
dose2 = rbind(c(1:5), c(0.10, 0.22, 0.41, 0.64, 0.81))
sim3p32 = sim3p3(dose2)
show_table(sim3p32,dose2)
# b
dose3 = rbind(c(1:5), c(0.043, 0.10,0.22, 0.41, 0.64))
sim3p33 = sim3p3(dose3)
show_table(sim3p33, dose3)
dose4 = rbind(c(1:5), c(0.007, 0.017, 0.043, 0.10, 0.22))
sim3p34 = sim3p3(dose4)
show_table(sim3p34, dose4)
dose5 = rbind(c(1:5), c(0.003, 0.007, 0.017, 0.043, 0.10))
sim3p35 = sim3p3(dose5)
show_table(sim3p35, dose5)
```

## Question 3

### a)

```{r}
show_CRM = function(sim, PI){
  res_table = data.frame(dose = c(1:5), trueDLT = PI, sp = sim$MTD, anTreat = sim$level, anDLT = sim$tox)
  res_table %>% 
    knitr::kable()
}
```

Perfomance metrics of 3+3 design and CRM with different does-toxicity curves are shown below.

```{r}
library(dfcrm)
set.seed(123)
p0 = c(0.08, 0.14, 0.25, 0.37, 0.52)
# 1
PI1 = c(0.017, 0.043, 0.010, 0.22, 0.41)
sim1 = crmsim(PI1, p0, 0.1, 31, x0=3, nsim = 1000, count = FALSE)
show_CRM(sim1, PI1)
show_table(sim3p31,dose1)
# 2
PI2 = c(0.10, 0.22, 0.41, 0.64, 0.81)
sim2 = crmsim(PI2, p0, 0.1, 31, x0=3, nsim = 1000, count = FALSE)
show_CRM(sim2, PI2)
show_table(sim3p32,dose2)
# 3
PI3 = c(0.043, 0.10, 0.22, 0.41, 0.64)
sim3 = crmsim(PI3, p0, 0.1, 31, x0=3, nsim = 1000, count = FALSE)
show_CRM(sim3, PI3)
show_table(sim3p33,dose3)
# 4
PI4 = c(0.007, 0.017, 0.043, 0.10, 0.22)
sim4 = crmsim(PI4, p0, 0.1, 31, x0=3, nsim = 1000, count = FALSE)
show_CRM(sim4, PI4)
show_table(sim3p34,dose4)
# 5
PI5 = c(0.003, 0.007, 0.017, 0.043, 0.10)
sim5 = crmsim(PI5, p0, 0.1, 31, x0=3, nsim = 1000, count = FALSE)
show_CRM(sim5, PI5)
show_table(sim3p35,dose5)
```

By comparing performance metrics of 3+3 design and CRM design we can find that

1) The selection probability of true MTD in CRM design is always the highest among 5 doses, while the dose with the highest selection probability in 3+3 design is always the dose near MTD.

2) Average numbers of patients treated in 3+3 design is similar among each dose while in CRM, more people treated with MTD.

3) Average numbers of patients with DLT in CRM overall are larger than those in 3+3 design.

### b)

```{r}
PCS = c(sim1$MTD[3], sim2$MTD[1], sim3$MTD[2], sim4$MTD[4], sim5$MTD[5])
mean(PCS)
```

The probability of correctly selecting(PCS) the MTD average over the five scenarios is `r mean(PCS)`.

