---
title: "Computing assignment 3"
author: "Xinyi Lin"
date: "11/3/2019"
output: pdf_document
---

*Write code to run simulation for a two-armed trial (A vs B) according to the followings:*

*Each subject will be evaluated for a binary response, with probability $p_A$ and $p_B$ respectively for arm A and arm B;*

* The total sample size is $N= 120$;

* The trial enrolls in 4 blocks each with $n= 30$ subjects in each block;

* In Block 1, subjects are randomized to arms A and B with equal probability.

* At the end of each block, update the randomization probabilities using Thompson’s rulewith no attenuation (i.e., set $a= 1$).

*1. Run 1,000 simulated trials under Scenario 1: $p_A=p_B= 0.5$.*

*(a) Perform a chi-squared test to compare A and B at the end of each trial. Record the proportion of trials with P-value less than 0.05.*

```{r}
simAR = function(pA=0.5, pB=0.5, N=120, n=30){
  # block 1
  assignment1 = rbinom(n,1,0.5) #1->A, 0->B
  resA = rbinom(sum(assignment1), 1, pA)
  resB = rbinom(n-sum(assignment1), 1, pB)
  #response and nonresponse numbers
  A1 = rep(NA, 4)
  A0 = rep(NA, 4)
  B1 = rep(NA, 4)
  B0 = rep(NA, 4)
  A1[1] = sum(resA)
  A0[1] = length(resA)-A1[1]
  B1[1] = sum(resB)
  B0[1] = length(resB)-B1[1]
  #pA, pB follows Beta(1,1)
  alphaA = 1
  betaA = 1
  alphaB = 1
  betaB = 1
  # block 2-4
  for (i in 2:4) {
    #calcualte rA
    alphaA = alphaA+sum(resA)
    betaA = betaA+length(resA)-sum(resA)
    alphaB = alphaB+sum(resB)
    betaB = betaB+length(resB)-sum(resB)
    sim_pA = rbeta(1000, alphaA, betaA)
    sim_pB = rbeta(1000, alphaB, betaB)
    rA = sum(sim_pA>sim_pB)/1000
    #get simulation results
    assignment = rbinom(n,1,rA)
    resA = rbinom(sum(assignment), 1, pA)
    resB = rbinom(n-sum(assignment), 1, pB)
    #response and nonresponse numbers
    A1[i] = sum(resA)
    A0[i] = length(resA)-A1[i]
    B1[i] = sum(resB)
    B0[i] = length(resB)-B1[i]
  }
  res = rbind(A1,A0,B1,B0)
  return(res)
}
```

```{r, warning=FALSE}
set.seed(123)
chi_res = rep(NA, 1000)
for (i in 1:1000) {
  res = simAR()
  sum_A1 = sum(res[1,])
  sum_A0 = sum(res[2,])
  sum_B1 = sum(res[3,])
  sum_B0 = sum(res[4,])
  sum_res = matrix(c(sum_A1, sum_B1, sum_A0, sum_B0), ncol = 2)
  chi_res[i] = chisq.test(sum_res)$p.value
}
mean(chi_res<0.05)
```

The proportion of trials with P-value less than 0.05 is 0.035.

*2. Run 1,000 simulated trials under Scenario 2: $p_A= 0.3$ and $p_B= 0.6$.*

*(a) Perform a chi-squared test to compare A and B at the end of the each trial. Record the proportion of trials with P-value less than 0.05.*

```{r, warning=FALSE}
set.seed(123)
chi_res = rep(NA, 1000)
for (i in 1:1000) {
  res = simAR(pA=0.3, pB=0.6)
  sum_A1 = sum(res[1,])
  sum_A0 = sum(res[2,])
  sum_B1 = sum(res[3,])
  sum_B0 = sum(res[4,])
  sum_res = matrix(c(sum_A1, sum_B1, sum_A0, sum_B0), ncol = 2)
  chi_res[i] = chisq.test(sum_res)$p.value
}
mean(chi_res<0.05)
```

The proportion of trials with P-value less than 0.05 is 0.677.

*(b) For a non-adaptive, 1:1 randomized two-armed trial withn= 60 in each arm, what is the power of a chi-squared test under Scenario B?*

Power is the propability to accepte alternative hypothesis when alternative hypothesis is ture. In scenairo 2, the null hypothesis is $p_A=p_B$ and the alternative hypothesis is ture. So the power is the probability to reject the null hypothesis.

```{r}
non_ada = function(n = 60, pA = 0.3, pB = 0.6){
  # arm A
  resA = rbinom(n, 1, pA)
  resB = rbinom(n, 1, pB)
  A1 = sum(resA)
  A0 = n-A1
  B1 = sum(resB)
  B0 = n-B1
  res = c(A1, B1, A0, B0)
  return(res)
}
```

```{r}
set.seed(123)
chi_res = rep(NA, 1000)
for (i in 1:1000) {
  res = non_ada()
  sum_res = matrix(res, ncol = 2)
  chi_res[i] = chisq.test(sum_res)$p.value
}
mean(chi_res<0.05)
```

*3. Perform Mantel Haenzel test to the data in Lecture 06 Adaptive randomization, slide #30.*

```{r}
data30 <- array(c(3, 1, 7, 9,
                   8, 1, 15, 6,
                   16, 2, 10, 2,
                   33, 1, 6, 0),
                 dim = c(2, 2, 4),
                 dimnames = list(ARM = c("ARM0", "ARM1"),
                                 Response = c("R1", "R0"),
                                 Block = 1:4))
mantelhaen.test(data30)
```

As p-value is 0.2852 which is larger than 0.05, we cannot reject the null hypothesis and conclude that there is no association between trials and outcomes.

*4. Run 1,000 simulated trials under Scenario 3: $pA=pB= 0.3$ in Blocks 1 and 2, and improved response rate $pA=pB= 0.6$ in Blocks 3 and 4.*

*(a) Perform a chi-squared test to compare A and B at the end of each trial. Record the proportion of trials with P-value less than 0.05.*

```{r}
simAR3 = function(N=120, n=30){
  # block 1
  assignment1 = rbinom(n,1,0.5) #1->A, 0->B
  resA = rbinom(sum(assignment1), 1, 0.3)
  resB = rbinom(n-sum(assignment1), 1, 0.3)
  #response and nonresponse numbers
  A1 = rep(NA, 4)
  A0 = rep(NA, 4)
  B1 = rep(NA, 4)
  B0 = rep(NA, 4)
  A1[1] = sum(resA)
  A0[1] = length(resA)-A1[1]
  B1[1] = sum(resB)
  B0[1] = length(resB)-B1[1]
  #pA, pB follows Beta(1,1)
  alphaA = 1
  betaA = 1
  alphaB = 1
  betaB = 1
  #calcualte rA
  alphaA = alphaA+sum(resA)
  betaA = betaA+length(resA)-sum(resA)
  alphaB = alphaB+sum(resB)
  betaB = betaB+length(resB)-sum(resB)
  sim_pA = rbeta(1000, alphaA, betaA)
  sim_pB = rbeta(1000, alphaB, betaB)
  rA = sum(sim_pA>sim_pB)/1000
  #get simulation results
  assignment = rbinom(n,1,rA)
  resA = rbinom(sum(assignment), 1, 0.3)
  resB = rbinom(n-sum(assignment), 1, 0.3)
  #response and nonresponse numbers
  A1[2] = sum(resA)
  A0[2] = length(resA)-A1[2]
  B1[2] = sum(resB)
  B0[2] = length(resB)-B1[2]
  # block 3-4
  for (i in 3:4) {
    #calcualte rA
    alphaA = alphaA+sum(resA)
    betaA = betaA+length(resA)-sum(resA)
    alphaB = alphaB+sum(resB)
    betaB = betaB+length(resB)-sum(resB)
    sim_pA = rbeta(1000, alphaA, betaA)
    sim_pB = rbeta(1000, alphaB, betaB)
    rA = sum(sim_pA>sim_pB)/1000
    #get simulation results
    assignment = rbinom(n,1,rA)
    resA = rbinom(sum(assignment), 1, 0.6)
    resB = rbinom(n-sum(assignment), 1, 0.6)
    #response and nonresponse numbers
    A1[i] = sum(resA)
    A0[i] = length(resA)-A1[i]
    B1[i] = sum(resB)
    B0[i] = length(resB)-B1[i]
  }
  res = rbind(A1,A0,B1,B0)
  return(res)
}
```

```{r, warning=FALSE}
set.seed(123)
chi_res = rep(NA, 1000)
for (i in 1:1000) {
  res = simAR3()
  sum_A1 = sum(res[1,])
  sum_A0 = sum(res[2,])
  sum_B1 = sum(res[3,])
  sum_B0 = sum(res[4,])
  sum_res = matrix(c(sum_A1, sum_B1, sum_A0, sum_B0), ncol = 2)
  chi_res[i] = chisq.test(sum_res)$p.value
}
mean(chi_res<0.05)
```

The proportion of trials with P-value less than 0.05 is 0.308.

*(b) Perform Mantel Haenzel test to compare A and B at the end of each trial. Record the proportion of trials with P-value less than 0.05.*

```{r}
set.seed(123)
ManH_res = rep(NA, 1000)
for (i in 1:1000) {
  res = simAR3()
  simAR3_data <- array(c(res[1,1], res[3,1], res[2,1], res[4,1],
                    res[1,2], res[3,2], res[2,2], res[4,2],
                    res[1,3], res[3,3], res[2,3], res[4,3],
                    res[1,4], res[3,4], res[2,4], res[4,4]),
                  dim = c(2, 2, 4),
                  dimnames = list(ARM = c("ARM0", "ARM1"),
                                  Response = c("R1", "R0"),
                                  Block = 1:4))
  ManH_res[i] = mantelhaen.test(simAR3_data)$p.value
}
mean(ManH_res<0.05)
```

The proportion of trials with P-value less than 0.05 in Mantel Haenzel test is 0.308.