---
title: "rstan"
author: "Xinyi Lin"
date: "10/28/2019"
output: html_document
---

```{r}
library(rstan)
library(tidyverse)
```

```{r}
options(mc.cores = parallel::detectCores())

rstan_options(auto_write = TRUE)
```

## Session 1

### part 1

Simulate data

```{r}
set.seed(1)
n = 30
beta0 = 10
beta1 = 5
sigma = 2
x = rnorm(n, mean = 10)
epsilon = rnorm(n)
y = beta0 + beta1*x + epsilon
plot(y~x)
```

Run stan model

```{r}
stan_data = list(N = 30, x = x, y = y)
stan_fit = stan(file = "./Bayesian Computation/1LinearRegression.stan", data = stan_data, warmup = 500, iter = 1000, chains = 4, seed = 1)
print(stan_fit)
posterior = rstan::extract(stan_fit)
```

fit `lm` function to test estimators get through stan.

```{r}
lm1.model = lm(y~x)
summary(lm1.model)

posterior_est = data.frame(iter = 1:2000, beta0 = posterior$beta0, beta1 = posterior$beta1, sigma = posterior$sigma) %>% 
  gather(key = "parameter", value = "value", beta0:sigma)
  

posterior_est %>% 
  ggplot(aes(x = iter, y = value, color = parameter)) +
  geom_line()

traceplot(stan_fit)
```

### part 2

Changing our priors

Prameters in .stan file for norm is different(second parameter in .stan is se)

```{r}
stan_data = list(N = 30, x = model.matrix(~x), y = y, K = 2) # K-dimension of predictors
stan_fit2 = stan(file = "./Bayesian Computation/2LinearRegression.stan", data = stan_data, warmup = 500, iter = 1000, chains = 4, seed = 1)
print(stan_fit2)
```

### Part 4

```{r}
stan_fit3 = stan(file = "./Bayesian Computation/3LinearRegression.stan", data = stan_data, warmup = 500, iter = 1000, chains = 4, seed = 1)
y_rep = as.matrix(stan_fit3, pars = "y_rep")
dim(y_rep)
```

## Session 2: Binomial outcome and hierarchical model

```{r}
y = c(6, 1, 10, 0, 11, 1, 0, 1, 11, 0, 6, 0, 9, 11, 25, 1)
S = c(17, 7, 25, 1, 20, 3, 3, 1, 27, 3, 16, 0, 23, 14, 39, 4)
Drug = rep(c("Erlotinib", "Erlotinib+bexaorotane", "Vandetanib", "Sorafenib"), each = 4)
Marker = rep(c("EGFR", "KRAS/BRAF", "VEGF", "RXR/Cyclin D1"), times = 4)
mydata = data.frame(y, S, Drug, Marker)
mydata1 = mydata[which(mydata$Marker == "EGFR"),]
```

### Part 1

Binomial 

```{r}
stan_data = list(N = 4, y = mydata1$y, S = mydata1$S)
stan_fit1a = stan(file = "./Bayesian Computation/1aBinomial.stan", data = stan_data,warmup = 500, iter = 1000, chains = 4, seed = 1)
print(stan_fit1a)
```

# Logistic regression model

```{r}
stan_data = list(N = 4, X = model.matrix(~Drug, data = mydata1), y = mydata1$y,S = mydata1$S, D = 4)
stan_fit1b = stan(file = "./Bayesian Computation/1bLogisticRegression.stan", data = stan_data,warmup = 500, iter = 1000, chains = 4, seed = 1)
print(stan_fit1b)
posterior1b = rstan::extract(stan_fit1b)
```

### Part 3:Binomial outcome with different prior

```{r}
stan_data = list(N = 4, X = model.matrix(~Drug, data = mydata1),y = mydata1$y, S = mydata1$S, D = 4)
stan_fit2 = stan(file = "./Bayesian Computation/2LogisticRegression.stan", data = stan_data,warmup = 500, iter = 1000, chains = 4, seed = 1)
print(stan_fit2)
```

### Part3: Hierarchical model

Binomial model

```{r}
mydata2 = mydata[which(mydata$Drug == "Erlotinib+bexaorotane"),]
stan_data = list(N = 4, y = mydata2$y, S = mydata2$S)
stan_fit3a = stan(file = "./Bayesian Computation/3aBinomialHierarchical.stan", data = stan_data,warmup = 500, iter = 1000, chains = 4, seed = 1)
print(stan_fit3a)
```

Logistic model

```{r}
stan_data = list(N = 16, X = model.matrix(~Drug, data = mydata), y = mydata$y,S = mydata$S, D = 4, m = rep(seq(1:4), times = 4), M = 4)
stan_fit3b = stan(file = "./Bayesian Computation/3bLogisticHierarchical.stan", data = stan_data,warmup = 500, iter = 1000, chains = 4, seed = 1)
print(stan_fit3b)
```

